\chapter{Theory}

This chapter attempts to build a theoretical foundation for the thesis by providing an understanding of significant developments in the field, different models, preprocessing techniques, and assessment measures. Its goal is to provide the reader with the necessary knowledge to interpret the research findings and their importance.

\section{Natural Language Processing}

In the domains of artificial intelligence and computer science, the study of the interaction between computers and human language is referred to as \acf{nlp} \citep{helland_tackling_2023}. It seeks to make it possible for machines to produce, interpret, and process human language. By employing a variety of techniques and different approaches, such as deep learning, rule-based systems, and statistical methods, is NLP capable of tackling different language-related tasks. The usage of it can be found in numerous areas, including machine translation, chatbots, text classification, and speech recognition \citep{helland_tackling_2023}.

\section{Machine Learning}

The term "machine learning" describes the creation of computer programmes that use past performance to solve problems and finish tasks. The measurement of the performance is calculated by the ability to do so \citep{helland_tackling_2023}. A training dataset represents the “experience” that is acquired by machine learning models, which contains output and input pairs. FThrough the examination of these instances, the model is able to identify and extrapolate the patterns to new, unobserved data \citep{helland_tackling_2023}. Simplifying, it can be understood as a reflection of how people learn and adjust to new situations. Fraud detection, self-driving cars, and personalized suggestions are examples of common applications \citep{helland_tackling_2023}.

\section{Deep Learning}

Deep learning is a method that uses non-linear modules to alter data at several levels of abstraction, allowing models to find patterns in the raw data \citep{helland_tackling_2023}. According to this, deep learning algorithms may be able to identify and acquire new data features without requiring human assistance \citep{helland_tackling_2023}. Because of its universal learning, generalization potential, robustness, and scalability advantages, this can be used in a variety of applications without the need for precise feature engineering \citep{helland_tackling_2023}.

\section{Evaluation Metrics}

A machine learning model's performance is assessed using metrics. They are employed to evaluate the accuracy of the made predictions, to analyze the output of various models, and to fine-tune them for optimal performance \citep{helland_tackling_2023}. Different types of machine learning issues have various types of metrics available. The model selection process, the optimization procedure, and the overall understanding of the model's capabilities can all be impacted by the metrics chosen. Incorrect metric selection may also lead to a biased model that is at contrast with the goals of the project \citep{helland_tackling_2023}.

\subsection{Accuracy}

One popular assessment metric for classification problems is “accuracy”. Out of all the samples in the prediction, it calculates the proportion of correctly classified samples \citep{helland_tackling_2023}. 

Accuracy = Number of correctly classified instances / Total number of instances

Since accuracy provides an true representation of the model's capabilities and performance, it is helpful when working with balanced datasets \citep{helland_tackling_2023}. Accuracy may be deceptive and more challenging to understand when datasets are imbalanced \citep{helland_tackling_2023}. This might be the result of the dataset having a single label that covers the majority of the samples. In this instance, a respectable degree of accuracy can still be obtained by projecting all samples to the dominating label. This does not suggest that the model is flawed because it does not take into account the less prevalent but equally important labels. Accuracy in multi-label classification only takes into account samples where every label is correctly classified. Because of this, using accuracy as a multi-label classification metric to evaluate the performance of multi-label models is more strict, less informative, and less desirable \citep{helland_tackling_2023}.